---
title: "BIS15L Final Project"
author: "Group 11"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: false
  pdf_document:
    toc: false
---


Data from "Meta-analysis of yield response of foliar fungicide-treated hybrid corn in the United States and Ontario, Canada," by Kiersten Wise et al. Can be accessed at https://pubmed.ncbi.nlm.nih.gov/31166965/



***Loading and cleaning libraries and data, establishing NAs, and selecting variables (Catalina):***


```{r message=FALSE, warning=FALSE}

#Loading libraries

library(tidyverse)
library(janitor)
library(shiny)
library(shinydashboard)
library(naniar)
library(readxl)
``` 


```{r}

#Reading data

Analysis <- read_excel("data/fungicide treated corn/Analysis.xlsx")

#Creating a table for the fungicide key for ShinyApp

key <- clean_names(read_excel("data/fungicide treated corn/Analysis.xlsx", sheet = "Notes")) %>% 
  as.data.frame(key) %>% 
  select(fungicide_codes) %>% 
  separate(fungicide_codes, c("code", "fungicide"), sep = "/") 

```

```{r}

#Cleaning, selecting, and filtering data.

corn <- clean_names(Analysis) %>% 
  select(location, year, state, rep, fung, timing, total_disease, yield_kg_ha, utc_total_disease, utc_yield_kg_ha, total_disease_control, yield_change_kg_ha) %>% 
  replace_with_na_all(condition~.x %in% c(".", common_na_numbers, common_na_strings)) %>% #Replacing NA replacements with real NAs, our data used "." instead of anything else but using common numbers and strings just in case
  mutate(
    repetition = as.numeric(rep), 
    fungicide = as.factor(fung),
    treated_disease_total = as.numeric(total_disease),
    treated_yield = as.numeric(yield_kg_ha),
    control_disease_total = as.numeric(utc_total_disease),
    control_yield = as.numeric(utc_yield_kg_ha),
    yield_change = as.numeric(yield_change_kg_ha)
  ) %>% #mutating into numerics and factors for graphing purposes
  select(
    location, state, year, repetition, fungicide, timing, treated_disease_total, treated_yield, control_disease_total, control_yield, yield_change
  ) %>% 
  filter(location != "GA") #filtering out Georgia because it had no values to compare with in 2014.
```



```{r}
names(corn)
```


**Variables:**

*location:* Location within a state.

*state:* US state/Canadian province where the crops were treated.

*year:* Year of treatment.

*fungicide:* Values 1-19, each corresponding to a fungicide and application style, definitions for each 'code' can be found in 'key' table. Corresponding also to "repetition" and "timing" values.

*treated_disease_total:* Total disease (fungi) in the treated crops.

*treated_yield:* Yield of crops produced by the treated group (kg/ha)

*control_disease_total:* Total disease (fungi) in control/untreated groups.

*control_yield:* Yield of crops produced by control/untreated groups (kg/ha).

*yield_change:* Difference in treated and untreated yields (kg/ha)




***Graphical Analysis (Alejandro)***

Comparing different fungicides effectiveness by State

```{r}
corn %>% 
  mutate(treated_disease_total=as.numeric(treated_disease_total),
         fung=as.factor(fungicide)) %>% 
  ggplot(aes(x=fung, y=treated_disease_total, fill=fung))+
  geom_col()+
  facet_wrap(~state) +
  labs(title = "Fungicide Effectiveness By State", y = "Disease Total (treated)", x = "Fungicide")

#MS, SD, and ND have almost no effective treatment across the board, and everyone else minus AR, GA, and LA, have at least one that doesnt work at all
```

How Fast can Fungus evolve higher resistance from one year to another?
```{r}
corn %>% 
  filter(year=="2014") %>% 
  mutate(treated_disease_total=as.numeric(treated_disease_total),
         fung=as.factor(fungicide)) %>% 
  ggplot(aes(x=fung, y=treated_disease_total, fill=fung))+
  geom_col()+
  facet_wrap(~state) +
  labs(title = "Amount of Treated Disease per State (2014)", x = "Fungicide", y = "Diseases Treated (count)")
```

```{r}
corn %>% 
  filter(state==c("AR", "IA", "IL", "IN", "LA", "ND", "Ontario", "WI")) %>%  #Filtered for these for less clutter, since these only ones in 2014,IL no show because no 2015 data
  filter(year=="2015") %>% 
  mutate(treated_disease_total=as.numeric(treated_disease_total),
         fungicide=as.factor(fungicide)) %>% 
  ggplot(aes(x=fungicide, y=treated_disease_total, fill=fungicide))+
  geom_col()+
  facet_wrap(~state) +
  labs(title = "Amount of Treated Disease per State (2015)", x = "Fungicide", y = "Diseases Treated (count)")
```

```{r}
names(corn)
```

Which Fungicide is the most effective at treating overall?
```{r}
corn %>% 
  mutate(treated_disease_total=as.numeric(treated_disease_total)) %>% 
  group_by(fungicide) %>% 
  summarize(total_treated=sum(treated_disease_total, na.rm=T))
```

```{r}
corn %>% 
  mutate(treated_disease_total=as.numeric(treated_disease_total),
          fungicide=as.factor(fungicide)) %>% 
  group_by(fungicide) %>% 
  summarize(total_treated=sum(treated_disease_total, na.rm=T)) %>% 
  ggplot(aes(x=fungicide, y=total_treated, fill=fungicide))+
  geom_col() +
  labs(title = "Total Amount of Disease Treated by Fungicide", x = "Fungicide", y = "Total Disease Treated (count)")
```


Which fungicide helps produce the most amount of corn?
```{r}
corn %>% 
mutate(treated_yield=as.numeric(treated_yield)) %>%
  group_by(fungicide) %>% 
  summarize(total_harvest=sum(treated_yield, na.rm = T))
```
```{r}
corn %>% 
mutate(treated_yield=as.numeric(treated_yield),
       fungicide=as.factor(fungicide)) %>%
  group_by(fungicide) %>% 
  summarize(total_harvest=sum(treated_yield, na.rm = T)) %>% 
  ggplot(aes(x=fungicide, y=total_harvest, fill=fungicide))+
  geom_col() +
  labs(title = "Corn Yield by Fungicide", x = "Fungicide", y = "Total Yield (kg/ha)")
```

```{r}
corn %>% 
mutate(treated_yield=as.numeric(treated_yield),
        fungicide=as.factor(fungicide)) %>%
  group_by(fungicide, state) %>% 
  summarize(total_harvest=sum(treated_yield, na.rm = T))
```


Does this change by state?
```{r}
corn %>% 
mutate(treated_yield=as.numeric(treated_yield),
        fungicide=as.factor(fungicide)) %>%
  group_by(fungicide, state) %>% 
  summarize(total_harvest=sum(treated_yield, na.rm = T)) %>% 
  ggplot(aes(x=state, y=total_harvest, fill=state))+
  geom_col()+
  facet_wrap(~fungicide) +
  labs(title = "Total Harvest by State per Fungicide", x = "State", y = "Treated Yield (kg/ha)")
```
```{r}
corn %>% 
mutate(treated_yield=as.numeric(treated_yield),
        fungicide=as.factor(fungicide)) %>%
  group_by(fungicide, state) %>% 
  summarize(total_harvest=sum(treated_yield, na.rm = T)) %>% 
  ggplot(aes(x=fungicide, y=total_harvest, fill=fungicide))+
  geom_col()+
  facet_wrap(~state) +
  labs(title = "Total Yield per Fungicide by State", x = "Fungicide", y = "Total Yield (kg/ha)")
```



Does Fungicide treatment make a noticeable difference?
```{r}
corn %>% 
  mutate(treated_yield=as.numeric(treated_yield),
         control_yield=as.numeric(control_yield)) %>%
  group_by(state) %>% 
  summarize(total_harvest_treated=sum(treated_yield, na.rm=T),
            total_harvest_untreated=sum(control_yield, na.rm=T),
            gross_diff_=total_harvest_treated-total_harvest_untreated)
```

```{r}
corn %>% 
  mutate(treated_yield=as.numeric(treated_yield),
         control_yield=as.numeric(control_yield)) %>%
  group_by(fungicide) %>% 
  summarize(total_harvest_treated=sum(treated_yield, na.rm=T),
            total_harvest_untreated=sum(control_yield, na.rm=T),
            gross_diff_=total_harvest_treated-total_harvest_untreated) %>% 
  ggplot(aes(x = fungicide, y = gross_diff_)) +
  geom_col(aes(fill = fungicide)) +
  labs(title = "Difference in Overall Treated and Untreated Yields per Fungicide", x = "Fungicide", y = "Difference (kg/ha)")
```

```{r}
corn %>% 
  mutate(yield_change=as.numeric(yield_change)) %>%
  group_by(fungicide)%>%
  ggplot(aes(x=fungicide, y=yield_change, fill=fungicide))+
  geom_boxplot() +
  labs(title = "Average Yield Difference in Treated and Untreated Crops by Fungicide", x = "Fungicide", y = "Change in Yield (kg/ha)")
```






***Shiny App Comparing these Variables (Vaishnavi)***

```{r, eval=F, echo=T}
library(shiny)
ui <- dashboardPage(
  
  dashboardHeader(title = "Corn Fungicide Yield Analysis"),
  
  dashboardSidebar(
    
    selectInput(
      "state_select",
      "Select a State:",
      choices = sort(unique(corn$state)),
      selected = unique(corn$state)[1]
    ),
    
   h4("Fungicide Key"),
    tableOutput("fungicide_key"),
    
    helpText("Select a state to explore fungicide performance and yield impact.")
    
  ),
  
  #Added Fungicide Key ^^
  
  dashboardBody(
    
    fluidRow(
      
      valueBoxOutput("top_fungicide"),
      valueBoxOutput("avg_yield"),
      valueBoxOutput("avg_control_yield")
      
    ),
    
    fluidRow(
      
      box(
        title = "Yield by Fungicide",
        status = "success",
        solidHeader = T,
        width = 6,
        plotOutput("yield_plot", height = 300)
      ),
      
      box(
        title = "Disease Levels by Fungicide",
        status = "warning",
        solidHeader = T,
        width = 6,
        plotOutput("disease_plot", height = 300)
      )
      
    ),
    
    
    fluidRow(
      
      box(
        title = "Yield Comparison (Treated vs Control)",
        status = "success",
        solidHeader = T,
        width = 12,
        plotOutput("comparison_plot", height = 350)
      )
      
    )
    
  )
)


server <- function(input, output, session) {
  
  state_data <- reactive({
    
    corn %>%
      filter(state == input$state_select) %>%
      mutate(
        treated_yield = as.numeric(treated_yield),
        control_yield = as.numeric(control_yield),
        treated_disease_total = as.numeric(treated_disease_total)
      )
    
  })
  
  
  #Fungicide with highest yield in selected state
  best_fungicide <- reactive({
    
    state_data() %>%
      group_by(fungicide) %>%
      summarize(avg_yield = mean(treated_yield, na.rm = T)) %>%
      arrange(desc(avg_yield)) %>%
      slice(1)
    
  })
  
  
  output$top_fungicide <- renderValueBox({
    
    valueBox(
      value = best_fungicide()$fungicide,
      subtitle = paste("Highest Yield Fungicide in", input$state_select),
      color = "purple"
    )
    
  })
  
  
  output$avg_yield <- renderValueBox({
    
    valueBox(
      round(mean(state_data()$treated_yield, na.rm = T), 2),
      "Average Treated Yield (kg/ha)",
      color = "red"
    )
    
  })
  
  
  output$avg_control_yield <- renderValueBox({
    
    valueBox(
      round(mean(state_data()$control_yield, na.rm = T), 2),
      "Average Untreated Yield (kg/ha)",
      color = "green"
    )
    
  })
  
  
  #Yield by fungicide plot
  output$yield_plot <- renderPlot({
    
    state_data() %>%
      ggplot(aes(x = factor(fungicide), y = treated_yield, fill = factor(fungicide))) +
      geom_boxplot(show.legend = F) +
      labs(
        x = "Fungicide",
        y = "Yield (kg/ha)",
        title = paste("Yield Distribution by Fungicide in", input$state_select)
      ) +
      theme_minimal(base_size = 14)
    
  })
  
  
  #Disease plot
  output$disease_plot <- renderPlot({
    
    state_data() %>%
      ggplot(aes(x = factor(fungicide), y = treated_disease_total, fill = factor(fungicide))) +
      geom_boxplot(show.legend = F) +
      labs(
        x = "Fungicide",
        y = "Disease Level",
        title = paste("Disease Levels by Fungicide in", input$state_select)
      ) +
      theme_minimal(base_size = 14)
    
  })
  
  
  #Treated vs untreated comparison
  output$comparison_plot <- renderPlot({
    
    state_data() %>%
      pivot_longer(
        cols = c(treated_yield, control_yield),
        names_to = "treatment",
        values_to = "yield"
      ) %>%
      ggplot(aes(x = treatment, y = yield, fill = treatment)) +
      geom_boxplot() +
      labs(
        x = "",
        y = "Yield (kg/ha)",
        title = paste("Yield Improvement from Fungicide in", input$state_select)
      ) +
      theme_minimal(base_size = 14)
    
  })
  
  
  #Fungicide Key Table
  output$fungicide_key <- renderTable({
    
    key
    
  })
  
}

shinyApp(ui, server)
```

