---
title: "BIS15L Final Project"
author: "Group 11"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: no
---


Loading Libraries: 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(shiny)
library(shinydashboard)
library(naniar)
``` 



Loading and cleaning data, establishing NAs, and selecting columns
```{r}
corn <- clean_names(Analysis) %>% 
  select(location, year, state, rep, fung, timing, total_disease, yield_kg_ha, utc_total_disease, utc_yield_kg_ha, total_disease_control) %>% 
  mutate(
    repetition = rep, 
    fungicide = fung,
    treated_disease_total = total_disease,
    treated_yield = yield_kg_ha,
    control_disease_total = utc_total_disease,
    control_yield = utc_yield_kg_ha,
    yield_change = total_disease_control
  ) %>% 
  select(
    location, state, year, repetition, fungicide, timing, treated_disease_total, treated_yield, control_disease_total, control_yield, yield_change
  ) %>% 
  filter(location != "GA" & location != "Ontario") %>% 
  replace_with_na_all(condition~.x %in% c(".", common_na_numbers, common_na_strings))
```


Comparing different fungicides effectiveness by State

```{r}
corn %>% 
  mutate(treated_disease_total=as.numeric(treated_disease_total),
         fung=as.factor(fungicide)) %>% 
  ggplot(aes(x=fung, y=treated_disease_total, fill=fung))+
  geom_col()+
  facet_wrap(~state)

#MS, SD, and ND have almost no effective treatment across the board, and everyone else minus AR, GA, and LA, have at least one that doesnt work at all
```

How Fast can Fungus evolve to resist from one year to another?
```{r}
corn %>% 
  filter(year=="2014") %>% 
  mutate(treated_disease_total=as.numeric(treated_disease_total),
         fung=as.factor(fungicide)) %>% 
  ggplot(aes(x=fung, y=treated_disease_total, fill=fung))+
  geom_col()+
  facet_wrap(~state)
```

```{r}
corn %>% 
  filter(state==c("AR", "IA", "IL", "IN", "LA", "ND", "Ontario", "WI")) %>%  #Filtered for these for less clutter, since these only ones in 2014,IL no show because no 2015 data
  filter(year=="2015") %>% 
  mutate(treated_disease_total=as.numeric(treated_disease_total),
         fungicide=as.factor(fungicide)) %>% 
  ggplot(aes(x=fungicide, y=treated_disease_total, fill=fungicide))+
  geom_col()+
  facet_wrap(~state)
```

```{r}
names(corn)
```

Which Fungicide is the most effective at treating overall?
```{r}
corn %>% 
  mutate(treated_disease_total=as.numeric(treated_disease_total)) %>% 
  group_by(fungicide) %>% 
  summarize(total_treated=sum(treated_disease_total, na.rm=T))
```

```{r}
corn %>% 
  mutate(treated_disease_total=as.numeric(treated_disease_total),
          fungicide=as.factor(fungicide)) %>% 
  group_by(fungicide) %>% 
  summarize(total_treated=sum(treated_disease_total, na.rm=T)) %>% 
  ggplot(aes(x=fungicide, y=total_treated, fill=fungicide))+
  geom_col()
```

Which one saves the most amount of corn?

```{r}
corn %>% 
mutate(treated_yield=as.numeric(treated_yield)) %>%
  group_by(fungicide) %>% 
  summarize(total_harvest=sum(treated_yield, na.rm = T))
```

Does this change by state?
```{r}
corn %>% 
mutate(treated_yield=as.numeric(treated_yield)) %>%
  group_by(fungicide, state) %>% 
  summarize(total_harvest=sum(treated_yield, na.rm = T)) #I'll finish later but going to make graphs to compare top and bottom, fungicide 9 is most effective overall but going to make graph to see where it works best and which states its beat by other
```


Does Fungicide treatment make a noticeable difference?
```{r}
corn %>% 
  mutate(treated_yield=as.numeric(treated_yield),
         control_yield=as.numeric(control_yield)) %>%
  group_by(state) %>% 
  summarize(total_harvest_treated=sum(treated_yield, na.rm=T),
            total_harvest_untreated=sum(control_yield, na.rm=T))
#Will also finish graph part later
```

